%------------------------
% Resume Template
% Author : Anubhav Singh
% Github : https://github.com/xprilion
% License : MIT
%------------------------

\documentclass[a4paper,20pt]{article}

\usepackage{latexsym}
\usepackage[empty]{fullpage}
\usepackage{titlesec}
\usepackage{marvosym}
\usepackage[usenames,dvipsnames]{color}
\usepackage{verbatim}
\usepackage{enumitem}
\usepackage[pdftex]{hyperref}
\usepackage{fancyhdr}

\pagestyle{fancy}
\fancyhf{} % clear all header and footer fields
\fancyfoot{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

% Adjust margins
\addtolength{\oddsidemargin}{-0.530in}
\addtolength{\evensidemargin}{-0.375in}
\addtolength{\textwidth}{1in}
\addtolength{\topmargin}{-.45in}
\addtolength{\textheight}{1in}

\urlstyle{rm}

\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

% Sections formatting
\titleformat{\section}{
  \vspace{-10pt}\scshape\raggedright\large
}{}{0em}{}[\color{black}\titlerule \vspace{-6pt}]

%-------------------------
% Custom commands
\newcommand{\resumeItem}[2]{
  \item\small{
    \textbf{#1}{: #2 \vspace{-2pt}}
  }
}

\newcommand{\resumeItemWithoutTitle}[1]{
  \item\small{
    {\vspace{-2pt}}
  }
}

\newcommand{\resumeSubheading}[4]{
  \vspace{-1pt}\item
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \textbf{#1} & #2 \\
      \textit{#3} & \textit{#4} \\
    \end{tabular*}\vspace{-5pt}
}


\newcommand{\resumeSubItem}[2]{\resumeItem{#1}{#2}\vspace{-3pt}}

\renewcommand{\labelitemii}{$\circ$}

\newcommand{\resumeSubHeadingListStart}{\begin{itemize}[leftmargin=*]}
\newcommand{\resumeSubHeadingListEnd}{\end{itemize}}
\newcommand{\resumeItemListStart}{\begin{itemize}}
\newcommand{\resumeItemListEnd}{\end{itemize}\vspace{-5pt}}

%-----------------------------
%%%%%%  CV STARTS HERE  %%%%%%

\begin{document}

%----------HEADING-----------------
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
  \textbf{{\LARGE Pratul Muthuraja}} & Email: \href{mailto:}{pratulmuthuraja@gmail.com}\\
  \href{https://pratulmuthuraja.com}{Portfolio: pratulmuthuraja.com} & Mobile:~~~+91-848-9468-800 \\
  \href{https://github.com/pratulmuthuraja}{Github: ~~github.com/pratulmuthuraja} \\
\end{tabular*}

%-----------EDUCATION-----------------
\section{Education}
  \resumeSubHeadingListStart
    \resumeSubheading
      {Illinois Insitute of Technology}{Chicago, USA}
      {Master of Computer Science}{January 2021 - May 2023}
      {\scriptsize \textit{ \footnotesize{\newline{}\textbf{Courses:} Big Data, Cloud Computing, Software Engineering, Virtual Machines}}}
    \resumeSubheading
      {Anna University}{Chennai, India}
      {Master of Computer Science}{June 2014 - May 2018}
      {\scriptsize \textit{ \footnotesize{\newline{}\textbf{Courses:} Operating Systems, Data Structures, Analysis Of Algorithms, Networking, Databases}}}
    \resumeSubHeadingListEnd
\vspace{-5pt}
\section{Skills}
	\resumeSubHeadingListStart
	\resumeSubItem{Languages}{~~~~~~Python, Go, JavaScript, SQL, Bash, HTML/CSS, Markdown}
	\resumeSubItem{Frameworks}{~~~~Django, Flask, ReactJS, NodeJS, LAMP}
	\resumeSubItem{Tools}{~~~~~~~~~~~~~~Kubernetes, Docker, GIT, PostgreSQL, MySQL, SQLite, NGINX, Helm, Gatsby, Hugo, Jenkins, ChatGPT}
	\resumeSubItem{Platforms}{~~~~~~~Linux, Web, Mac, AWS, GCP}
	\resumeSubItem{Soft Skills}{~~~~~~~Leadership, Event Management, Writing, Public Speaking, Time Management}

\resumeSubHeadingListEnd
\vspace{-5pt}
\section{Experience}
  \resumeSubHeadingListStart
    \resumeSubheading{Armour Chapter of Triangle}{Chicago, USA}
    {Full-Stack Developer (Volunteer, Full-time)}{November 2023 - November 2024}
    \resumeItemListStart
        \resumeItem{Tech Stack}{React.js, Nodejs, Javascript, TailwindCSS, PostgreSQL.}
        \resumeItem{Website}
          {Built a full-stack web application to securely manage social and event data for over 200K members.}
        \resumeItem{Optimize Backend}
          {Overhauled backend architecture, optimizing database queries and caching strategies to bring API response times under 100ms while cutting operational costs by 60\%.}
        \resumeItem{Workflow Efficiency}
          {Designed and implemented event-driven batch jobs for real-time notifications and automated task scheduling, improving engagement and workflow efficiency.}
        \resumeItem{User Interface}
          {Led UI/UX development, working closely with stakeholders to create Figma wireframes and translating them into a highly responsive React frontend.}
        \resumeItem{System Reliability}
          {Improved system reliability to 99.99\% uptime by setting up load balancing, auto-scaling, failover mechanisms, and database replication, minimizing downtime and ensuring high availability.}
        \resumeItem{Observability}
          {Introduced observability tools (logs, metrics, and traces) to enable proactive monitoring and reduce incident resolution time by 40\%.}
      \resumeItemListEnd
\vspace{-5pt}
    \resumeSubheading
		{Illinois Institute of Technology}{Chicago, USA}
		{Graduate Student Ambassador (Part-time)}{August 2021 -  December 2022}
		\resumeItemListStart
        \resumeItem{Data Analysis}
          {Managed databases and generated reports on Salesforce; Analyzed prospective student data to improve response
to Graduate Discover Day event resulting in 19\% growth of incoming graduate students compared to 2020.}
        \resumeItem{Leadership}
          {Strong leadership and collaboration skills evidenced by organizing recruitment activities through campus tours,
student panel, and events for the Graduate Admissions Office showing 37\% better response rate.}
		\resumeItemListEnd

    \resumeSubheading
		{Narayana Glass Company}{Madurai, India}
		{Business Development Manager (Full-time)}{June 2018 -  December 2020}
		\resumeItemListStart
        \resumeItem{Manager}
          {Upgraded heavy machinery resulting in increased productivity by 25\% and implemented better safety standards.}
        \resumeItem{Market Penetration}
          {Achieved national market penetration resulting in 15\% sales increase and 10\% market value increase.}
		\resumeItemListEnd

\resumeSubHeadingListEnd

%-----------PROJECTS-----------------
\vspace{-5pt}
\section{Projects}
\resumeSubHeadingListStart
\resumeSubItem{Database Performance Analysis Tool}{(Graduate Research Project) Built an automated tool to deploy Cassandra, MongoDB and CouchDB clusters on AWS utilizing Terraform and Python scripts, resulting in a reduction in deployment time by 75\%. Conducted automated rigorous testing in 2, 4, and 8 cluster modes to evaluate performance based on pre-defined
criteria and recorded results using Matplotlib graphs for Data Visualization. Tech: Python, Cassandra, MongoDB, CouchDB, Terraform.}
\vspace{2pt}
\resumeSubItem{Home Lab} {Designed and built a 2 node cluster running Proxmox VE to learn, test, and experiment with self-hosting and server management. Honed automation skills and achieved 50\% reduced reliance on cloud services. Implemented logging and monitoring using Loki and Grafana. Used Open Media Vault to convert HDD array to NAS. Set up remote access through Cloudflared Tunnel and used NGINX proxy manager to act as a reverse proxy. Tech Stack: Proxmox VE, Debian, Docker, Kubernetes, Bash, Loki, Grafana, Cloudflared Tunnel, NGINX.}
\vspace{2pt}
\resumeSubItem{Portfolio Website}{Built portfolio website using Hugo, an opensource static site generator considering its fast build times, vast plugins library, speed, ease of integration, and security to showcase technical skills, experience, personal projects,
and achievements.  Tech: Hugo, Go, Markdown.}
\vspace{2pt}
\resumeSubItem{Open-source Contributor - Meshery}{Have been contributing to Meshery, an open-source, cloud-native management plane designed to help with the deployment, operation, and lifecycle management of service meshes and cloud-native infrastructure since March 2025. Tech: Gatsby, React.js, Go.}
\vspace{2pt}
\resumeSubItem{Search Engine and Social Network(Web Development, Web Crawler, Search)}{Created from scratch a social network and a search engine based on the idea of integrating Facebook and Google. The launched website was among top 1000 websites in India during 2012-2013. Tech: PHP, MySQL, HTML, CSS, WebSockets, JavaScript, RSS, XML ( May '12)}
\resumeSubHeadingListEnd
\vspace{-5pt}

% \section{Publications}
% \resumeSubHeadingListStart
% \resumeSubItem{Book: Deep Learning on Web (Web Development, Deep Learning)}{Work in Progress book to be published by Packt Publishing in late 2019. Tech: Django, Python, AWS, GCP, Azure (November '18)}
% \vspace{2pt}
% \resumeSubItem{Book: Deep Learning on Mobile Devices (Flutter App Development, Deep Learning)}{Work in Progress book to be published by Packt Publishing in late 2019. Tech: Flutter, Android, Firebase, TensorFlow, Python, Dart (December '18)}
% \resumeSubHeadingListEnd

% \vspace{-5pt}
% \section{Volunteer Experience}
%   \resumeSubHeadingListStart
% 	\resumeSubheading
%     {Community Lead at Developer Student Clubs NSEC}{Kolkata, India}
%     {Conducted online and offline technical \& soft-skills training impacting over 3000 students.}{Jan 2019 - Present}
% \vspace{5pt}
%     % \vspace{10pt}\textbf{\large{Community Experience}}
%     \resumeSubheading
%     {Event Organizer at Google Developers Group Kolkata}{Kolkata, India}
%     {Organized events, conducted workshops and delivered workshops reaching over 7000 developers.}{Jan 2018 - Present}
% \resumeSubHeadingListEnd

\end{document}
